% Options for packages loaded elsewhere
% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
]{article}
\usepackage{xcolor}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother


\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother





\setlength{\emergencystretch}{3em} % prevent overfull lines

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}



 


\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\@ifpackageloaded{fontawesome5}{}{\usepackage{fontawesome5}}
\definecolor{quarto-callout-color}{HTML}{909090}
\definecolor{quarto-callout-note-color}{HTML}{0758E5}
\definecolor{quarto-callout-important-color}{HTML}{CC1914}
\definecolor{quarto-callout-warning-color}{HTML}{EB9113}
\definecolor{quarto-callout-tip-color}{HTML}{00A047}
\definecolor{quarto-callout-caution-color}{HTML}{FC5300}
\definecolor{quarto-callout-color-frame}{HTML}{acacac}
\definecolor{quarto-callout-note-color-frame}{HTML}{4582ec}
\definecolor{quarto-callout-important-color-frame}{HTML}{d9534f}
\definecolor{quarto-callout-warning-color-frame}{HTML}{f0ad4e}
\definecolor{quarto-callout-tip-color-frame}{HTML}{02b875}
\definecolor{quarto-callout-caution-color-frame}{HTML}{fd7e14}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\usepackage{amsthm}
\theoremstyle{definition}
\newtheorem{exercise}{Exercise}[section]
\theoremstyle{remark}
\AtBeginDocument{\renewcommand*{\proofname}{Proof}}
\newtheorem*{remark}{Remark}
\newtheorem*{solution}{Solution}
\newtheorem{refremark}{Remark}[section]
\newtheorem{refsolution}{Solution}[section]
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Math 557 August 29},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}


\title{Math 557 August 29}
\author{}
\date{}
\begin{document}
\maketitle


\subsection{Key Concepts}\label{key-concepts}

\begin{itemize}
\item
  \textbf{\(\mathcal{L}\)-structures}: Provide a realm to interpret
  \(\mathcal{L}\)-formulas over.

  \begin{itemize}
  \tightlist
  \item
    non-empty universe
  \item
    interpretation of all constant, function, and relation symbols
  \end{itemize}
\item
  \textbf{structure isomorphism}: bijection \(\pi\) between the
  universes of two \(\mathcal{L}\)-structures that preserves
  interpretations of symbols. General version of the well-known idea of
  isomorphism of mathematical structures.
\item
  \textbf{assignment}: function \(\alpha: \mathcal{V} \to M\) that
  assigns all variables values in the universe of a structure.
  Recursively extends to terms.
\item
  \textbf{satisfaction}: \(\mathcal{M} \models \varphi[\alpha]\) means
  the \(\mathcal{L}\)-formula \(\varphi\) is true when interpreted in
  the \(\mathcal{L}\)-structure \(\mathcal{M}\) and all variables are
  assigned values according to assignment \(\alpha\). This is defined
  recursively over the structure of a formula.
\end{itemize}

\section{Problems}\label{problems}

\begin{exercise}[]\protect\hypertarget{exr-}{}\label{exr-}

\hfill\break
The set of \(\cal{L}\)-terms \(\mathcal{T}^{\mathcal{L}}\) is defined as
the \emph{smallest} set (with respect to \(\subseteq\)) containing
variables and constant symbols and that is closed under function symbol
application.

\begin{itemize}
\item
  Why do this set exist?
\item
  Does a similar argument work for the set of \(\mathcal{L}\)-formulas?
\item
  Find an example of a (non-empty) family of sets and a property \(P\)
  (applicable to the sets in the family) such that a
  \(\subseteq\)-smallest set with property \(P\) does not exist.
\end{itemize}

\end{exercise}

\begin{exercise}[]\protect\hypertarget{exr-}{}\label{exr-}

\hfill\break
Prove the \textbf{Readbility Lemma}: For any term \(t\), exactly one of
the following holds.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  There exists a \(v \in \mathcal{V}\) such that \(t \equiv v\).
\item
  There exists a \(c \in \mathcal{C}^{\mathcal{L}}\) such that
  \(t \equiv c\).
\item
  There exists a number \(n \in \mathbb{N}\), a function symbol \(f
  \in F_n^{\mathcal{L}}\), and terms \(t_1,
  \dots, t_n\) such that \(t \equiv ft_1\dots t_n\).
\end{enumerate}

\end{exercise}

\begin{exercise}[]\protect\hypertarget{exr-}{}\label{exr-}

\hfill\break
Show that no proper initial segment of a term is a term. (\emph{Use
induction on the length of a term.})

Deduce \textbf{unique readability}: The decomposition in (3.) above is
unique.

\end{exercise}

\begin{exercise}[]\protect\hypertarget{exr-}{}\label{exr-}

\hfill\break
Let \(n\) be a natural number and let \(+L\) be any language. Find
formulas \(, \psi^{\geq}, \psi^{\leq}, \psi^=\) such that, for any
\(+L\)-structure \(\mathcal{M}\),

\begin{align}
\mathcal{M} \models \psi^{\geq} & \iff \; |M| \geq n \\
\mathcal{M} \models \psi^{\leq} & \iff \; |M| \leq n \\
\mathcal{M} \models \psi^{=} & \iff \; |M| = n \\
\end{align}

\end{exercise}

\begin{exercise}[]\protect\hypertarget{exr-}{}\label{exr-}

\hfill\break
Let \(+L= \{\cdot , e\}\) be the language of groups. Show that there is
a sentence \(\varphi\) such that \(\mathcal{M} \models \varphi\) if and
only if \(M \cong \mathbb{Z}/2\mathbb{Z}\times \mathbb{Z}/2\mathbb{Z}\).

\end{exercise}

\begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, toprule=.15mm, arc=.35mm, coltitle=black, opacitybacktitle=0.6, opacityback=0, breakable, colframe=quarto-callout-important-color-frame, bottomrule=.15mm, toptitle=1mm, title=\textcolor{quarto-callout-important-color}{\faExclamation}\hspace{0.5em}{Take-home Problem}, bottomtitle=1mm, left=2mm, colbacktitle=quarto-callout-important-color!10!white, titlerule=0mm, leftrule=.75mm, colback=white]

\hfill\break
Scott sentences for finite languages.

\end{tcolorbox}




\end{document}
